---
title: "Assignment B1"
author: "Katia Rosenflanz"
date: "2024-10-28"
output: github_document
---

```{r}
library(palmerpenguins)
library(datateachr)
library(tidyverse)
library(testthat)
```


```{r}
#' Calculate Summary Statistics by Group
#' 
#' @description 
#' The function calculates four summary statistics (min, max, mean, standard deviation) of a variable
#' by group within a data table. The variables used for grouping and summarizing are input by the user.
#' 
#' @param data Input data.  
#' @param sort_var Variable used for grouping; named as sort_var to indicate that it is used to sort 
#' the data into groups for subsequent analysis 
#' @param stat_var Numeric variable which is used to calculate summary statistics. Named stat_var as
#' this is the variable with which the summary statistics are calculated 
#'
#' @return A tibble is returned which contains the grouping variable and summary statistics as columns (min, max, mean, and standard deviation) 
sum_stats <- function(data, sort_var, stat_var) {
  #check to see if the variable for summary statistics calculations is numeric 
  if (!is.numeric(pull(data, !!enquo(stat_var)))) {
    stop("The variable you are trying to summarize is not numeric.")
  }
  new_data <- {{data}} %>%
    group_by({{sort_var}}) %>%
    summarize(min = if(all(is.na({{stat_var}}))) NA_real_   #if all values are NA, treat NA as double/numeric missing value
              else min({{stat_var}}, na.rm = TRUE),         #remove all NA values before calculation
              max = if(all(is.na({{stat_var}}))) NA_real_   #if all values are NA, treat NA as double/numeric missing value
              else max({{stat_var}}, na.rm = TRUE), 
              mean= mean({{stat_var}}, na.rm = TRUE), 
              stdev = sd({{stat_var}}, na.rm = TRUE))
  return(new_data)
}
```

After the function has been defined, several examples can be used to demonstrate its function. 
The expected output is a tibble containing the grouped variable and the four summary statistics columns. 
```{r}
#Example 1: penguins dataset
#group by island, summarize the body_mass_g variable
sum_stats(penguins, island, body_mass_g)

```
```{r}
#Example 2: cancer_sample dataset
#group by diagnosis (benign or malignant), summarize the area_mean variable
sum_stats(cancer_sample, diagnosis, area_mean)

```
Next, I will test a use of the sum_stats function that I know will produce an error: calculating summary statistics on a variable which is non-numeric. In this case, the sum_stats function will try to summarize the root_barrier variable, which is not numeric; this will cause an error. 
```{r error = TRUE}
#Example 3: vancouver_trees dataset
#group by genus, calculate statistics on root barrier (which is nonnumeric)
sum_stats(vancouver_trees, genus_name, root_barrier)

```
Finally, I can run several formal tests on the function using the testthat package. 
```{r}
#### Testing the function ####
#test case 1: vector with no NAs
test_that("Sum_stats testing with data with no NAs works", {
  data_noNA <- data.frame(
    sort_var = c("A", "A", "B", "B", "C", "C"),
    stat_var = c(1,2,3,4,5,6)
  )
  expect_equal(sum_stats(data_noNA, sort_var, stat_var)$min, c(1, 3, 5))
})

#test case 2: vector with NAs 
test_that("Sum_stats testing with data including NAs should also work", {

  data_withNA <- data.frame(
    sort_var = c("A", "A", "B", "B", "C", "C"),
    stat_var = c(1,2, NA, NA, NA, 6)
  )
  expect_equal(sum_stats(data_withNA, sort_var, stat_var)$mean, c(1.5, NA, 6))
})

#test case 3: Data has non-numeric input as the variable for summarizing
test_that("Sum_stats fails with non-numeric input as summarizing variable", {
  
  data_nonNumeric <- data.frame(
    sort_var = c("A", "A", "B", "B", "C", "C"),
    stat_var = c("Hello", "Yes", "Mountain", "River", "Forest", "Desert")
  )
  expect_error(sum_stats(data_nonNumeric))
})
```



